---
title: "Sales Quantity Prediction Based on Advertising Media Investment"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

### Introduction:
Data was analyzed for a company that invested in television, radio, and 
newspaper advertising. The scope of this project was to provide an 
overview of the amounts that were spent in these different media and also to 
use machine learning to predict the quantity sold of the company's product 
based on their advertising investment. 
The advertising was done across 200 markets. The advertising expense was 
provided in terms of thousands of dollars. The quantity of product sold was 
provided in terms of thousands of units.

**Data Source:** Advertising dataset from [An Introduction to Statistical 
Learning](https://www.statlearning.com/resources-first-edition).


**Business Problem:** To analyze how advertising media affected the quantity sold 
of a product across markets. The goal of the project was to provide the 
company with insights on how investment in television, radio, and newspaper 
advertising influenced sales of their product.

#### The project was done using R and it consisted of two parts:  
* **Part 1:** An Exploratory Data Analysis was performed in order to gain an
understanding of how television, radio, and newspaper advertising 
affected the sales of the company's product across different markets. An 
overview of the advertising expense was presented to support the company with
planning and business strategy.

* **Part 2:** The Multiple Linear Regression machine learning algorithm was used to 
build models to predict the company's sales, in terms of quantity of product
sold, based on the amounts that the company had invested in 
the advertising media. The models were then evaluated using the Adjusted 
R-squared metric and the Root Mean Square Error (RMSE) metric to select the 
best model to provide the sales prediction. The company could then use the
model to estimate the quantity of product they could expect to sell given the 
amounts that they planned to invest in the advertising media.

<br>

### Loading the Packages:
```{r Loading the Packages, message=FALSE}
library(tidyverse)
library(corrplot)
library(reshape2)
library(scales)
```

<br>

### Loading the Data:
```{r Loading the Data, message=FALSE}
# To load the data and to save it to a data frame called advertising_df:
advertising_df <- read_csv("Advertising.csv")
```

<br>

### Viewing the Data:
```{r Viewing the Data, message=FALSE}
# To view the first six rows of the data frame:
head(advertising_df)
```

The independent variables were TV, radio, and newspaper. The dependent variable was sales.

<br>

The first column of the data frame was an index indicating the number assigned
to each specific market. This column was removed, because it was not used in 
the analysis.
```{r}
# To remove the first column of the data frame:
advertising_df <- advertising_df %>% select(TV, radio, newspaper, sales)
```

<br>

The "radio", "newspaper", and "sales" columns were then capitalized.
```{r}
# To rename the "radio", "newspaper", and "sales" columns:
advertising_df <- advertising_df %>% 
  rename(Radio = radio, Newspaper = newspaper, Sales = sales)
```

<br>

The advertising expense values for television, radio, and newspaper were 
provided in terms of thousands of dollars. In order to change the formatting,
the values in each column were multiplied by 1,000.
```{r}
# To first create a copy of the original data frame:
advertising_df_v1 <- advertising_df

# To multiply the "TV" column by 1,000:
advertising_df_v1$TV <- advertising_df_v1$TV * 1000

# To multiply the "Radio" column by 1,000:
advertising_df_v1$Radio <- advertising_df_v1$Radio * 1000

# To multiply the "Newspaper" column by 1,000:
advertising_df_v1$Newspaper <- advertising_df_v1$Newspaper * 1000
```

<br>

The values for quantity of product sold were provided in terms of thousands of 
units. In order to change the formatting, these values were multiplied 
by 1,000.
```{r}
# To multiply the "Sales" column by 1,000:
advertising_df_v1$Sales <- advertising_df_v1$Sales * 1000
```

<br>

### Viewing the Structure of the Data:
```{r Viewing the Structure of the Data}
glimpse(advertising_df_v1)
```

<br>

### Summary Statistics:
```{r}
summary(advertising_df_v1)
```

#### **Findings:**
* Television was the most expensive form of advertising for the company.
* Radio was the least expensive form of advertising for the company.

<br>

The data frame was examined to determine if there were any duplicate rows.
```{r}
# To examine the data frame for any duplicate rows:
nrow(advertising_df_v1[duplicated(advertising_df_v1), ])
```

There were no duplicate rows found in the data frame.

<br>

The data frame was inspected for any missing values.
```{r}
# To inspect the data frame for missing values:
sapply(advertising_df_v1,function(x) sum(is.na(x)))
```

There were no missing values in the data frame.

<br>

## **Part 1:**
### Exploratory Data Analysis:
A box plot was created to analyze the amount that was spent on advertising 
for each media:
```{r, message=FALSE}
# To remove the "Sales" column when creating the box plot:
box_plot_df <- advertising_df_v1 %>% select(-Sales)

# To reshape the box_plot_df data frame to the long format:
box_plot_df_long <- melt(box_plot_df)

# To create the box plot:
ggplot(box_plot_df_long, aes(variable, value)) + 
  geom_boxplot() +
  scale_y_continuous(labels=scales::dollar_format(), limits = c(0, 300000), 
                     breaks = seq(0, 300000, by = 50000)) +
  labs(title = "Advertising Expense by Media",
       x = "Media",
       y = "Advertising Expense") +
  theme(text = element_text(size = 12.50))
```

#### **Findings:**
* Television had the greatest range in the amount spent by the company for advertising across their markets.
* Advertising on television was considerably more expensive than the other two media.
* The newspaper advertising expense and radio advertising expense resulted in ranges that were similar to each other, with the exception of two newspaper 
outlier values.

<br>

A numerical summary of the box plot was created.
```{r}
# To create a numerical summary of the box plot:
summary(box_plot_df)
```

#### **Findings:**
* The highest advertising expense for the company was television.
* The lowest advertising expense for the company was radio.
* The median amount spent on **TV** advertising was $149,750, followed by 
$25,750 spent on **Newspaper** advertising, and $22,900 spent on **Radio**
advertising. The mean values and the third quartile values displayed the 
same trend.

<br>

The outlier values for newspaper advertising were examined.
```{r}
# To examine the outlier values for newspaper advertising:
boxplot(advertising_df_v1$Newspaper, plot=FALSE)$out
```

The newspaper advertising expense outlier values were $114,000 and $100,900. 
These high values could have been due to a data entry error, or it was just 
very expensive for the company to advertise in newspapers in those two 
particular markets.

<br>

A box plot was created to check if the "Sales" column had any outliers.
```{r}
# To create a box plot to check for outliers in the "Sales" column:
ggplot(aes(x = "", y = Sales), data = advertising_df_v1) + 
  geom_boxplot() +
  scale_y_continuous(labels = scales::comma, limits = c(0, 30000)) +
  labs(x = "",
       y = "Quantity of Product Sold",
       title = "Box Plot of Sales") +
  theme(text = element_text(size = 12.25))
```

There were no outliers present in the "Sales" column.

<br>

A numerical summary was created for the "Sales" column box plot.
```{r}
# To create a numerical summary of the box plot for the "Sales" column:
advertising_df_v1 %>% select(Sales) %>% summary()
```

The lowest quantity of units sold was 1,600 units and the highest quantity of 
units sold was 27,000 units. The average (mean) quantity of units sold was
14,022 units.

<br>

A bar chart was then created to analyze the total amount that the company 
spent on each advertising media.
```{r}
# To create a bar chart to analyze the total amount that the company spent on 
# each advertising media:
box_plot_df_long %>%
  group_by(variable) %>%
  summarize(sum = sum(value, na.rm=TRUE)) %>%
  ungroup() %>%
  ggplot(aes(reorder(factor(variable), -sum), sum)) + 
  geom_col(width = 0.4) +
  geom_text(aes(label = dollar(sum),
                vjust = -0.7, hjust = 0.5)) +
  labs(title = "Total Expense by Advertising Media",
       x = "Advertising Media",
       y = "Total Expense") +
  theme(text = element_text(size = 12.50)) +
  scale_y_continuous(labels=scales::dollar_format(), limits = c(0, 32000000), 
                   breaks = seq(0, 32000000, by = 8000000))
```

The company invested most of its advertising budget in television, followed by
newspaper, and then radio.

<br>

A pie chart was created to analyze the budget allocation percentage for the 
advertising media.
```{r}
# To create a pie chart of the budget allocation:
box_plot_df_long %>%
  group_by(variable) %>%
  summarize(sum = sum(value, na.rm=TRUE)) %>%
  ungroup() %>% 
  mutate(prop = sum/sum(sum), percent = round((prop*100), 2)) %>%
  ggplot(aes(x = "", y = percent, fill = variable)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  geom_text(aes(label = paste0(percent, "%")), size = 7.0, vjust = -0.5, 
            position = position_stack(vjust = 0.8)) +
  labs(x = NULL, y = NULL, fill = "Media") +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(title = "Percentage of Advertising Budget by Media",
       fill = "Media") +
theme(text = element_text(size = 20.00))
```

Approximately three quarters of the advertising budget was allocated to 
television. The percentage of the budget invested in newspaper advertising and 
radio advertising were approximately equal to each other.

<br>

A correlation matrix was created to analyze the relationships between the 
advertising media and the quantity of product sold.
```{r}
# To create a correlation matrix to analyze the relationships between the 
# variables:
round(cor(advertising_df_v1), 2)

corrplot(cor(advertising_df_v1))
```

#### **Findings:**
* Sales had a strong positive correlation with television advertising.
* Sales had a moderate positive correlation with radio advertising.
* Sales had a very weak positive correlation with newspaper advertising.
* There was some multicollinearity present between radio advertising and 
newspaper advertising. Multicollinearity occurs when independent variables are
correlated with each other, which can make it difficult to determine the 
effect of each independent variable on the dependent variable. To address
this, a solution is to remove one of the correlated independent variables and 
to use the one that has the strongest correlation with the dependent variable.

<br>

A scatter plot was created to visualize the relationship between sales and the 
amount spent on television advertising.
```{r, message=FALSE}
# Quantity of Product Sold vs. TV Advertising Expense:
ggplot(advertising_df_v1, aes(x = TV, y = Sales))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(x = "TV Advertising Expense",
       y = "Quantity of Product Sold",
       title = "Quantity of Product Sold vs. TV Advertising Expense") +
  theme(text = element_text(size = 12.25)) +
  scale_y_continuous(labels = scales::comma, limits = c(0, 28000), 
                     breaks = seq(0, 28000, by = 7000)) + #divided by 4
  scale_x_continuous(labels=scales::dollar_format(), limits = c(0, 300000),
                     breaks = seq(0, 300000, by = 60000)) # divided by 5
```

There was a strong positive correlation between the quantity of 
product sold and the amount spent on television advertising for the product.

<br>

A scatter plot was created to visualize the relationship between sales and the 
amount spent on radio advertising.
```{r, message=FALSE}
# Quantity of Product Sold vs. Radio Advertising Expense:
ggplot(advertising_df_v1, aes(x = Radio, y = Sales))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(x = "Radio Advertising Expense",
       y = "Quantity of Product Sold",
       title = "Quantity of Product Sold vs. Radio Advertising Expense") +
  theme(text = element_text(size = 12.25)) +
  scale_y_continuous(labels = scales::comma, limits = c(0, 28000), 
                     breaks = seq(0, 28000, by = 7000)) +
  scale_x_continuous(labels=scales::dollar_format(), limits = c(0, 50000),
                     breaks = seq(0, 50000, by = 10000))
```

There was a moderate positive correlation between the quantity of 
product sold and the amount spent on radio advertising for the product.

<br>

A scatter plot was created to visualize the relationship between sales and the 
amount spent on newspaper advertising.
```{r, message=FALSE}
# Quantity of Product Sold vs. Newspaper Advertising Expense:
ggplot(advertising_df_v1, aes(x = Newspaper, y = Sales))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  labs(x = "Newspaper Advertising Expense",
       y = "Quantity of Product Sold",
       title = "Quantity of Product Sold vs. Newspaper Advertising Expense") +
  theme(text = element_text(size = 12.25)) +
  scale_y_continuous(labels = scales::comma, limits = c(0, 28000), 
                     breaks = seq(0, 28000, by = 7000)) +
  scale_x_continuous(labels=scales::dollar_format(), limits = c(0, 115000),
                     breaks = seq(0, 115000, by = 23000))
```

There was a very weak positive correlation between the quantity of 
product sold and the amount spent on newspaper advertising for the product.

<br>

## **Part 2:**
### Multiple Linear Regression Machine Learning Models:
In machine learning, the independent variables are also referred to as
"features" and the dependent variable is also called the "target variable".

Multiple Linear Regression is a machine learning algorithm that is used to 
analyze the relationship between features and the target variable. It uses 
several features to predict the outcome of the target variable.

For this project, the Multiple Linear Regression machine learning algorithm 
was used to build models to predict the company's product sales quantity given
the different amounts that the company had invested in television, radio, and 
newspaper advertising. 

Two Multiple Linear Regression machine learning models were built and 
compared. The models were then evaluated by using the Adjusted R-squared
metric and the Root Mean Square Error (RMSE) metric.
The model with the highest Adjusted R-squared and the lowest RMSE was then 
selected to provide the company with a sales quantity prediction that was 
based on the amount invested in advertising media.

<br>

To begin the model building process, the data frame was first split into a 
training set and a testing set.
```{r}
# To split the data frame into a training set and a testing set:
library(caTools)

# To make the example reproducible:
set.seed(49)

split <- sample.split(advertising_df_v1$Sales, SplitRatio = 0.8)
training_data <- subset(advertising_df_v1, split == "TRUE")
testing_data <- subset(advertising_df_v1, split == "FALSE")
```

The training set was used for constructing the model.
The testing set was used for validating the model and for making predictions.
Model validation involves using a testing set to evaluate how well a model
fits data that was not used in training the model.
For this project, 80% of the data was used to generate the model and the 
remaining 20% of the data was used to test the model.

<br>

From the Exploratory Data Analysis, it was observed that multicollinearity was
present between the Radio and Newspaper features. The VIF (Variance Inflation 
Factor) metric was used to measure the degree of multicollinearity between the 
features.
```{r, message=FALSE}
# To test for multicollinearity by using VIF (Variance Inflation Factor):
library(car)
set.seed(49)
vif_test <- lm(Sales~., data = training_data)
vif(vif_test)
```

#### **Findings:**
* VIF measures the increase in the variance of an estimated regression 
coefficient that is due to multicollinearity.
Multicollinearity occurs when two features are highly correlated with each 
other. If the degree of correlation is high enough between them, it may cause 
problems when fitting and interpreting the regression model. 

* A VIF value greater than five indicates possible severe correlation between 
one feature and another feature.
In this project, the VIF values for the Radio and Newspaper features were
above one, indicating that there was moderate multicollinearity present. 

 * To address the multicollinearity, two Multiple Linear Regression models were built and compared. The first model used all the features. The second model did not use the Newspaper feature. 
 
 * The Newspaper feature was not used in the second model, because the Exploratory Data Analysis showed that the correlation between Sales and the 
Radio feature was stronger than the correlation between Sales and the Newspaper feature.

<br>

The first Multiple Linear Regression model was built by using all the
features.
```{r}
# To make the example reproducible:
set.seed(49)
# To build the first model, using all the features:
model1 <- lm(Sales~. , data = training_data)
```

<br>

The summary of the first model was analyzed.
```{r}
# To view the summary of the initial model:
summary(model1)
```

#### Model Summary Interpretation:
##### **p-value:** 
* A p-value was calculated for each feature that was used to build the model. In
a Multiple Linear Regression model, if a feature has a p-value that is 
less than 0.05, this indicates that the feature has a statistically 
significant relationship with the target variable.

* The p-value for the TV feature was less than 0.05, indicating that there was
a statistically significant relationship between the amount invested in 
television advertising and the quantity of product that was sold.

* The p-value for the Radio feature was less than 0.05, indicating that there 
was a statistically significant relationship between the amount invested in 
radio advertising and the quantity of product that was sold. 

* The Newspaper feature did not have a statistically significant 
relationship with Sales, as its p-value was greater than 0.05. This supported
the findings from the Exploratory Data Analysis, where the Quantity of Product
Sold vs. Newspaper Advertising Expense scatter plot showed that there was a 
very weak positive correlation between the sales and the amount spent on 
newspaper advertising for the product.

##### **Multiple R-squared:** 
* Multiple R-squared is the coefficient of determination for a Multiple Linear
Regression model. It indicates how well the model fits the data by describing
the amount of variance in the target variable that can be explained by the 
features. 

* There is an important consideration with regards to Multiple 
R-squared. Adding more features to the model will increase Multiple R-squared,
even if the additional features do not add to the predicting power of the 
model. For this reason, the Adjusted R-squared should be used instead when 
comparing Multiple Linear Regression models.

##### **Adjusted R-squared:** 
* The Adjusted R-squared accounts for features that are not significant in the 
model. It indicates whether adding features improves a regression model 
or not. It only increases if the newly added features improve the model’s 
predicting power.

* The Adjusted R-squared also indicates how well the model fits the data. It 
describes the proportion of the variance in the target variable that can be 
explained by the features.
* In this model, the Adjusted R-squared value indicated that the features were 
able to explain 89.39% of the variance in the Sales target variable.

##### **F-statistic p-value:** 
* The F-statistic tests the significance of the overall model. It indicates
whether the model provides a better fit to the data than a model that contains
no features. Specifically, the p-value for the F-statistic indicates if the 
overall model fit is statistically significant. If the p-value for the 
F-statistic is less than 0.05, this indicates that there is a statistically 
significant relationship between the overall model and the target variable.

* In this model, the F-statistic p-value was less than the 0.05 significance 
level, indicating that the overall model fit was statistically 
significant. There was a statistically significant relationship between the 
overall model and Sales.

<br>

The Root Mean Square Error (RMSE) metric was then calculated by using the 
model on the testing set. In Multiple Linear Regression, the RMSE is used 
to determine how well a model fits the data. The RMSE is the standard 
deviation of the residuals. Residuals are the differences between the actual 
data values and the values that were predicted by the model. 
As the RMSE value decreases, a model is able to better fit a dataset.
The RMSE was calculated on the testing set in order to evaluate how well the 
model performed with data that was not used to train the model.
```{r}
# To calculate the RMSE for the testing set:
sqrt(mean((testing_data$Sales - predict(model1, newdata = testing_data))^2))
```

The testing set RMSE for this model was 1,500.6 units sold.
The RMSE was rounded to 1,501 units sold.

<br>

A second Multiple Linear Regression model was built, using only the TV and 
Radio features.
```{r}
# To make the example reproducible:
set.seed(49)
# To build the second model:
model2 <- lm(Sales~ TV + Radio, data = training_data)
```

<br>

The summary of the second model was analyzed.
```{r}
# To view the summary of the second model:
summary(model2)
```

#### Model Summary Interpretation:
##### **p-value:** 
* The p-value for the TV feature was less than 0.05, indicating that there was
a statistically significant relationship between the amount invested in 
television advertising and the quantity of product that was sold.

* The p-value for the Radio feature was less than 0.05, indicating that there 
was a statistically significant relationship between the amount invested in 
radio advertising and the quantity of product that was sold.

##### **Adjusted R-squared:** 
* The Adjusted R-squared indicated that the features were able to explain 89.45% 
of the variance in the Sales target variable. 
* Removing the Newspaper feature
increased the Adjusted R-squared from 89.39% to 89.45%, indicating that the 
second model was able to explain more of the variance in the target variable.

##### **F-statistic p-value:** 
* The F-statistic p-value was less than the 0.05 significance level, indicating
that the overall model fit was statistically significant. 
* There was a 
statistically significant relationship between the overall model and Sales.

<br>

The RMSE was then calculated for the testing dataset:
```{r}
# To calculate the RMSE for the testing dataset:
sqrt(mean((testing_data$Sales - predict(model2, newdata = testing_data))^2))
```

The testing set RMSE for this model was 1,488.663 units sold.
The RMSE was rounded to 1,489 units sold.

<br>

### Comparison of Multiple Linear Regression Models:
**Adjusted R-squared for the first model:** 89.39% 

**Testing set RMSE for the first model:**  1,501

<br>

**Adjusted R-squared for the second model:** 89.45%  

**Testing set RMSE for the second model:**  1,489  

<br>

When comparing several Multiple Linear Regression models, a higher Adjusted 
R-squared value is preferred. The Adjusted R-squared was higher in the second
model, indicating that the second model was able to explain more of the 
variance in the target variable. The second model was able to fit the data
better than the first model.

When comparing several Multiple Linear Regression models, a lower testing set 
RMSE value is preferred. The testing set RMSE was lower in the second model,
indicating that the second model was able to generate more accurate 
predictions than the first model.

<br>

**Conclusion:**
The second model was selected to provide the company with predictions of the
quantity of units sold based on advertising investment, because the second 
model had a higher Adjusted R-squared and a lower testing set RMSE than the 
first model. 

<br>


### Using the Selected Model to Make Predictions with New Data:
New data was created in order to use the second model to generate a
prediction for the quantity of units sold. The investment in television 
advertising was selected to be $3,000 and the investment in radio advertising 
was selected to be $2,000.
```{r}
# To use the second model to make a prediction on new data:
predict(model2, newdata = data.frame(TV = 3000, Radio = 2000))
```

If the company invested $3,000 in television advertising and $2,000 in radio
advertising, the model predicted that the company could expect to sell an
average of 3,453 units of their product.


### Key Takeaways:
* Television was the most expensive form of advertising for the company.

* Radio was the least expensive form of advertising for the company.

* Television had the greatest range in the amount spent by the company for 
advertising across their markets.

* Approximately three quarters of the advertising budget was allocated to 
television.

* The lowest quantity of units sold was 1,600 units and the highest quantity of
units sold was 27,000 units. The average (mean) quantity of units sold was 
14,022 units.

* The best performing Multiple Linear Regression machine learning model was 
selected to provide the company with predictions of the quantity of units sold 
based on advertising investment. The criteria for model selection was based on
which model had the highest Adjusted R-squared value and the lowest testing 
set RMSE value.

<br>

### Recommendations:
* Since sales were very weakly correlated with newspaper advertising, 
and newspaper advertising was more expensive than radio advertising, it is 
recommended that the company focuses on using the television and radio media.

* The company can use the selected machine learning model to obtain an estimate 
of the expected quantity of units sold, based on the amount that was budgeted 
for the advertising media.

<br>
<br>
<br>
<br>
<br>
<br>